---
title: "Fair game problems and answers"
output:
  html_document:
    df_print: paged
---

Consider a fair game which costs $Y$ to play: a fair coin is flipped twice, and the number of heads $X$ recorded. The game pays the player $X$.

1. Calculate $\mathbb{E}(X)$.

Here, $\mathbb{E}(X) = 0 . 1/4 + 1 . 1/2 + 2 . 1/4 = 1$

2. What is a fair price to play? Here, fair means that a player and the house should be even in the long-run.

The expected value of the game is the long-term mean, so a fair price is $1$.

3. Advanced: suppose you play the game 10 times. What's the probability that your profit exceeds your loss if $Y=1$? Check your answer in R.

We create a random variable $Z_t$ to indicate the profit after $t$ plays:

\begin{equation}
Z_t = \sum_{\tau=1}^{t} (X_t - 1) = \sum_{\tau=1}^{t} X_t - t
\end{equation}

where $X_t$ is a two-flip binomial with probability 1/2. Since the individual random variables all have the same probability, their sum is also a binomial (you can prove this via moment generating functions: see chapter in Blitzstein and Hwang book):

\begin{equation}
\zeta_t := \sum_{\tau=1}^{t} X_t
\end{equation}

follows: $\zeta_t \sim \mathcal{B}(2t, 0.5)$.

For 10 plays, the probability is then:

\begin{equation}
Pr(Z_{10} > 0) = Pr(\zeta_{10} > 10) = 1 - F_Z(10)
\end{equation}

which we can calculate using:

```{r}
1 - pbinom(10, 2 * 10, 0.5)
```

Checking via R
```{r}
one_game <- function() {
  X <- rbinom(1, 2, 0.5)
  X - 1
}

many_games <- function(t) {
  profit <- 0
  for(tau in 1:t)
    profit <- profit + one_game()
  profit
}

# now simulate many games
nreplicates <- 100000
profits <- vector(length = nreplicates)
t <- 10
for(i in 1:nreplicates)
  profits[i] <- many_games(t)
mean(profits > 0)
```

4. More advanced: suppose you start with a wealth of $1$ and continuing to replay the game until your profits fall to zero, at which point you stop. What is the probability that you last for 10 games or more?


Note, that the distribution is very long-tailed. So sometimes simulations of the game if left unchecked can last a long time. Here, since we are only interested in whether the number of plays exceeds a threshold, we don't directly simulate the game: instead we just determine whether this threshold is exceeded.
```{r}
one_game_revenue <- function() {
  X <- rbinom(1, 2, 0.5)
  X
}

many_games_exceeds_threshold <- function(threshold) {
  num_plays <- 0
  wealth <- 1
  while(wealth > 0 & num_plays < threshold) {
    wealth <- wealth + one_game_revenue() - 1
    num_plays <- num_plays + 1
  }
  ifelse(num_plays >= threshold, 1, 0)
}

# replicate many times
nreplicates <- 100000
num_plays <- vector(length = nreplicates)
for(i in 1:nreplicates)
  num_plays[i] <- many_games_exceeds_threshold(threshold=10)
mean(num_plays)
```

